---
title: "ATDT Experiment Data dB Trends"
author: "Lowe Wilsson"
date: '2022-06-03'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This script is for analyzing Auditory Tone Detection Task experiment data.

## Load necessary libraries
Install any missing libraries before running this.
```{r, warning=FALSE, message=FALSE}
library(ggplot2)
library(dplyr)
```

## Load data
Set current working directory to where this notebook is located and specify (relative) path to the directory which holds all of the experiment data CSV files.

```{r}
directory_this_document <- dirname(rstudioapi::getSourceEditorContext()$path)
setwd(directory_this_document)
DATA_DIR_PATH <- '../data'
```

Get a vector of all the CSV file names.

```{r}
data_filenames <- list.files(DATA_DIR_PATH)
# use regular expression to only keep CSV file names
data_filenames <- grep(".csv", data_filenames, value=TRUE)
```

Extract the participant id's from the filenames (here assuming that ID's do __not__ include any underscores) and combine them with timestamps to form participant 'codes' that will be used as 'keys' in list formed further down.

```{r}
part_ids <- gsub('([^_]*)_.*', '\\1', data_filenames)
part_codes <- gsub('([^_]*).+pure_tone_detection_(.*).csv', '\\1_\\2', data_filenames)
```

Keep only data files and participant codes from experiment runs where a non-blank participant ID was specified.

```{r}
nonblank_bool <- part_ids != ""
part_codes <- part_codes[nonblank_bool]
data_filenames <- data_filenames[nonblank_bool]
```

Load all of the data, putting each set in a data frame and collecting all data frames in a list.

```{r}
df_list <- list()
# loop over all data filenames
for (i in 1:length(data_filenames)) {
  # form full (directory path + filename) file path
  data_path <- file.path(DATA_DIR_PATH, data_filenames[i])
  part_id <- part_codes[i]
  df_list[[part_id]] <- read.csv(data_path, stringsAsFactors = FALSE)
}
```

## Collect all participants' dB data in a single 'long' data frame
Form a single data frame which holds _all_ 'desired dB' values, trial numbers, and participant ID's.

```{r}
# extract the dB and 'trial number' values from only 
# the rows of each data frame that have non-NA 'Desired_dB' values
# and are of 'Test' trialType
extract_nonna_db <- function(df) {
  db_vals <- df[['Desired_dB']]
  trial_vals <- df[['Trial']]
  ttype_vals <-  df[['trialType']]
  nonvalidtrial_bool <- is.na(db_vals) | is.na(trial_vals) | (ttype_vals != 'Test')
  db_vals <- db_vals[! nonvalidtrial_bool]
  trial_vals <- trial_vals[! nonvalidtrial_bool]
  return(list(db=db_vals, trial_num=trial_vals))
}
valdb_trialno_list <- lapply(df_list, extract_nonna_db)

# form a vector of all participants' trial 'desired dB' values
all_db_vals <- Reduce(
  function(x, y) c(x, y),
  lapply(valdb_trialno_list, function(x) {x[['db']]})
)
# form a vector of all participants' trial numbers
all_trial_nums <- Reduce(
  function(x, y) c(x, y),
  lapply(valdb_trialno_list, function(x) {x[['trial_num']]})
)

# form a vector which describes how many trials each participant went through
trials_per_participant <- sapply(
  lapply(valdb_trialno_list, function(x) {x[['trial_num']]}),
  max
)

# form a vector of participant id's, where each id is repeated based
# on the number of trials that the participant went through
all_pids <- rep(part_codes, times=trials_per_participant)

db_trial_df <- data.frame(
  participant_id=all_pids,
  trial_num=all_trial_nums,
  db=all_db_vals
)
```

## Visualize all participants' dB trends
Draw a plot which describes _all_ participants' 'desired dB' trends.

```{r}
ggplot(db_trial_df, aes(x=trial_num, y=db, color=participant_id)) +
  geom_line() +
  guides(color="none")
```

## Exclude participants with too few trials
Get the ID's of all participants who had 20 (limit arbitrarily chosen) or fewer trials.
```{r}
toofew_ids <- part_codes[trials_per_participant <= 20]
toofew_ids
```

Exclude participants with too few trials from further analysis.

```{r}
db_trial_gte30_df <- db_trial_df[!(db_trial_df$participant_id %in% toofew_ids), ]
```

Redraw the plot, excluding participants with too few trials.

```{r}
ggplot(db_trial_gte30_df, aes(x=trial_num, y=db, color=participant_id)) +
  geom_line() +
  guides(color="none")
```

## Summary statistics
The last trials' dB values should be fairly 'stable' and give reasonable estimates of participants' actual auditory thresholds. Thus, for each participant, use only the last 5 trials' dB values to calculate their auditory threshold. Then, calculate _group_ mean auditory threshold and standard deviation (based on each participant's 'mean-of-last-5-values').

```{r}
summary_df <- db_trial_gte30_df %>%
  group_by(participant_id) %>%
  slice_tail(n=5) %>%
  summarise(db_mean_participant=mean(db)) %>%
  ungroup() %>%
  summarise(
    db_mean=mean(db_mean_participant, na.rm=TRUE),
    db_sd=sd(db_mean_participant, na.rm=TRUE),
    n_participants=n()
  )
summary_df
```

The mean auditory threshold then was `r round(summary_df %>% pull('db_mean'), 2)` (SD=`r round(summary_df %>% pull('db_sd'), 2)`), based on data from `r summary_df %>% pull('n_participants')` participants.

We can do the same analysis again, but excluding outliers who have thresholds above 30dB, indicating that they probably didn't perform the experiment as expected and/or the experiment, e.g. participant answered randomly and/or audio wasn't correctly played.

```{r}
summary_df_sansoutliers <- db_trial_gte30_df %>%
  group_by(participant_id) %>%
  slice_tail(n=10) %>%
  summarise(db_mean_participant=mean(db)) %>%
  filter(db_mean_participant <= 30) %>%
  ungroup() %>%
  summarise(
    db_mean=mean(db_mean_participant, na.rm=TRUE),
    db_sd=sd(db_mean_participant, na.rm=TRUE),
    n_participants=n()
  )
summary_df_sansoutliers
```

